\vspace{-1in} \usefont{OT1}{bch}{b}{n} \normalfont \normalsize \textsc{CSEC-472 Authentication Paper} \\ [14pt] \huge A Security Analysis of the Lenel BlueDiamond \\

### 2020-05-05 paper/paper/paper.tex
Removing novel techniques
Added timeline and procedures

\begin{document}
\maketitle
\selectlanguage{english}
\begin{abstract}
Contactless authentication devices have existed in some capacity since the late 20th century. These devices allow users to authenticate themselves using something they have (smart card, proximity card, mobile device, etc) rather than something they know (passcode, combination, security question, etc) making the end-user less responsible for their authentication. This puts noticeably more pressure on the manufacturers to provide a product capable of providing confidentiality, integrity, and availability for its users. As such, the technology continues to advance as companies come up with newer, more advanced, ways to authenticate securely. In recent years, Lenel corporations has released multiple products under the BlueDiamond name that claim to "enhance freedom of movement in the workplace." /cite{lenelbluediamondwebsite} To date, these devices remain untested by the further research community for vulnerabilities that may lead to a breach of access control. As such, this paper aims to design an experimental process that future researchers may follow to accurately assess the security characteristics of Lenel's BlueDiamond contactless readers. This experiment will follow standard scientific procedure, using multiple tests to verify results and standard units of measurement to score success. Our analysis of these devices concludes that further testing on the subjects of Wiegand, RFID, Bluetooth, and the physical hardware may uncover significant findings.
\end{abstract}

@@

\section{Background \& Significance}
This  section  should  describe  relevant  security  concepts  related  to  your  topic  and address  historical  work  on  the  subject.  You  should  assume  that  the  reader  has  a reasonable amount of technical knowledge but is not an expert.

@@

It can be seen in the literature on BLE is very vulnerable to man-in-the-middle attacks, denial of service, packet sniffing, and other vulnerabilities. Even though higher levels of security are included in newer BLE versions, such as ECC or out-of-band key exchanges, security settings like PINs are still used by manufacturers \cite{8622000}\cite{jaihc19}\cite{mryan13}. A basic attack chain has emerged in the literature where the GATT is copied and an attack is launched to clear the device's Bluetooth pairings \cite{jaihc19}. The device is then tricked into pairing with the malicious device. The handshake is sniffed by the attacker allowing for encryption to be broken.

@@

\subsection{Procedures}
Words

\subsection{Timeline}
Words

\subsection{Novel Techniques}
Words

@@

\section{Acknowledgements}
We would like to thank Lenel for access to their BlueDiamond reader and their permission to conduct a test of the reader's security posture.

>>>>

\begin{document}
\maketitle
\selectlanguage{english}
\begin{abstract}
Contactless authentication devices have existed in some capacity since the late 20th century. These devices allow users to authenticate themselves using something they have (smart card, proximity card, mobile device, etc) rather than something they know (passcode, combination, security question, etc) making the end-user less responsible for their authentication. This puts noticeably more pressure on the manufacturers to provide a product capable of providing confidentiality, integrity, and availability for its users. As such, the technology continues to advance as companies come up with newer, more advanced, ways to authenticate securely. In recent years, Lenel corporations has released multiple products under the BlueDiamond name that claim to "enhance freedom of movement in the workplace." \cite{lenelbluediamondwebsite} To date, these devices remain untested by the further research community for vulnerabilities that may lead to a breach of access control. As such, this paper aims to design an experimental process that future researchers may follow to accurately assess the security characteristics of Lenel's BlueDiamond contactless readers. This experiment will follow standard scientific procedure, using multiple tests to verify results and standard units of measurement to score success. Our analysis of these devices concludes that further testing on the subjects of Wiegand, RFID, Bluetooth, and the physical hardware may uncover significant findings.
\end{abstract}

@@

\section{Background \& Significance}
Words

@@

It can be seen in the literature on BLE is very vulnerable to man-in-the-middle attacks, denial of service, packet sniffing, and other attacks. Even though higher levels of security are included in newer BLE versions, such as ECC or out-of-band key exchanges, security settings like PINs are still used by manufacturers \cite{8622000}\cite{jaihc19}\cite{mryan13}. A basic attack chain has emerged in the literature where the GATT is copied and an attack is launched to clear the device's Bluetooth pairings \cite{jaihc19}. The device is then tricked into pairing with the malicious device. The handshake is sniffed by the attacker allowing for encryption to be broken.

@@

\subsection{Procedures}
The testing performed against the readers would be broken into three primary categories: Bluetooth, RFID, and Wiegand.  These categories will follow logical attack chains to test a variety of vulnerabilites.

The Bluetooth procedures follow a primary attack chain to test for vulnerabilities.  The chain starts with either gattacker \cite{GattackingJasek} \cite{BluetoothMITMTal} or CrackLE \cite{10.1145/3190645.3190693} to gain an initial foothold.  From there, the next step is to test for either injection attacks or injection free attacks \cite{jaihc19} to clear the binding list.  After the binding list is cleared a man-in-the-middle attack can be attempted \cite{lonzetta2018security} \cite{osullivan} \cite{BluetoothMITMTal}.  From this point there are multiple attacks that can be attempted.  The first attack is to attempt to break the long-term key by brute forcing the term key \cite{sevier2019securityofble} \cite{179196}.  After this replay attacks and active sniffing can be attempted \cite{filizzola2018security} \cite{osullivan}.  Finally, the protocol could be reverse engineered to try and find vulnerabilities in the implementation.

The primary RFID attack is cloning, it is possible to try to prevent cloning though it is very hard \cite{7888545} \cite{dijkPortableBumping} \cite{mitrokotsa09} \cite{7945583}.  Secondly, the random number generator used for RFID cards can be tested for vulnerabilities \cite{6565237} \cite{mitrokotsa09}. A third possible set of attacks is to try passive and active relay against the reader \cite{hancke2009confidence} \cite{6810722}.  Another possibility is to examine the RFID card reader for possible injection attacks and remote code execution \cite{garciaIClass}.  The final attack method would be to attack the RFID chain of trust used in the reader \cite{lehtonen07}.

The Wiegand protocol attacks are relatively simple.  The first step of the attack chain is to analyze the various Wiegand outputs of the reader.  After a large enough scans have been made interception and duplication attacks can be made \cite{chung2017wiegand}. Finally, a BLEKey \cite{baseggio2015BLEKey} or similar device can be made and tested against the reader.  As a part of the Wiegand testing, attacks against the optical tamper sensor will need to be created and tested.

These three sets of attacks will cover a large amount of the reader's attack surface.  The particular kill chains are ordered to effectively build off of previous steps and to test the easier attacks before the more complicated ones.  The order also is set so that the possible dangerous attacks that could damage the device are run last.

\subsection{Timeline}
The timeline for this project assumes a eight week period for testing.  This eight week period will be broken up into five separate stages for setup, testing, and wrapup.  Eight weeks should be more than adequate to perform the necessary testing and provide a complete report.  The first period is one week long and dedicated to setup.  During this time a test lab would be built with the necessary equipment to test RFID, Bluetooth, and Wiegand.  The second period is two weeks in length and dedicated to testing RFID.  The third period is also two weeks in length, this is dedicated to testing for Bluetooth weaknesses.  The fourth period is the final two week period, this time dedicated to testing for Wiegand vulnerabilities.  The fifth and final period is one week in length, unlike the others this period is not for testing.  The last period is dedicated to writing the report to submit to Carrier for their review.

@@

\section{Acknowledgements}
We would like to thank Carrier for access to their BlueDiamond reader and their permission to conduct a test of the reader's security posture.
